---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
dt <- read.csv("activity.csv", na.strings = "NA")
dt$date <- as.Date(dt$date,"%Y-%m-%d")
dtomit <- na.omit(dt)
dt2 <- aggregate(steps ~ date, FUN = sum, data = dt)
meansteps <- mean(dt2$steps)
mediansteps <- median(dt2$steps)
```



## What is mean total number of steps taken per day?

```{r}
hist(dt2$steps, xlab="Steps", ylab = "Freq", main = "Mean total number of steps taken per day")
```

The mean of total number of steps taken per day is `r meansteps`.
The median of total number of steps taken per day is `r mediansteps`.

## What is the average daily activity pattern?

```{r}
dt3 <- aggregate(steps ~ interval, FUN = mean, data = dtomit)
plot(dt3, type = "l", xlab = "Interval", ylab = "Steps", main = "The average daily activity pattern")
```
The 5-minute interval contains the maximum number of step is `r dt3$interval[which.max(dt3$steps)]`.

## Imputing missing values
```{r}
naval <- sum(is.na(dt$steps))
```
Total number of missing values in the dataset is `r naval`

Make a new dataset with the missing data filled in with mean value of same interval
```{r}
dtnew <- dt
dtnew$steps[is.na(dt$steps)] <- dt3$steps[match(dt$interval, dt3$interval)][is.na(dt$steps)]
```

The new plot of new data
```{r}
dt2new <- aggregate(steps ~ date, FUN = sum, data = dtnew)
hist(dt2new$steps, xlab="Steps", ylab = "Freq", main = "Mean total number of steps taken per day in the new data")
```

## Are there differences in activity patterns between weekdays and weekends?

Make a new factor variable in the dataset with two levels -- "weekday" and "weekend"
```{r}
dtnew <- transform(dtnew, weekday = ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
```
Processing data for plotting
```{r}
dtweekday <- dtnew[dtnew$weekday == "Weekday",]
dtweekend <- dtnew[dtnew$weekday == "Weekend",]

dtweekday <- aggregate(steps ~ interval + weekday, FUN = mean, data = dtweekday)
dtweekend <- aggregate(steps ~ interval + weekday, FUN = mean, data = dtweekend)

dtwd <- rbind(dtweekend, dtweekday)
```
The plot of steps between weekend and weekday:
```{r}
library(ggplot2)
g <- qplot(interval, steps, data = dtwd, facets = .~weekday)
g + geom_line()
```